---
title: "Outcome-oriented educational tasks with flipbooks"
subtitle: "For *Chance Magazine's* 'Taking a chance in the classroom'"
author: "Evangeline Reynolds, June 2020"
output: pdf_document
  # xaringan::moon_reader:
  #   lib_dir: libs
  #   css: [default, hygge, ninjutsu]
  #   nature:
  #     ratio: 16:10
  #     highlightStyle: github
  #     highlightLines: true
  #     countIncrementalSlides: false
---




```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, 
                      comment = "", cache = F, fig.retina = 3, eval = T)
library(flipbookr)
library(tidyverse)
```





---

A goal of data science and statistics courses is giving students practice preparing and analyzing data.  A new style of assignment presents outcome-oriented exercises in 'flipbooks' (evamaerey.github.io/flipbooks/about).  Typically 'flipbooks' present code step-by-step and side-by-side with output. This gives you a movie-like experience linking new code to the change that it triggers in output.  But flipbookr can also be used in an *only output* mode.  This mode can be used in data science and statistics classrooms to create engaging outcome-oriented assignments!  

# What's this new style of exercise?

A flipbook data manipulation assignment might look as follows.  First, you might have a reference section of functions that can be used to accomplish the required tasks. Then a target data frame is introduced.  The target will be a final output of a data manipulation pipeline, where the input is a familiar data set (I use the 'gapminder' data set which my students having worked with for data visualization). Then, a walk-through from input to target --- including intermediary states --- are displayed.  Students are instructed to 'wrangle to it'; i.e. use the new data manipulation functions to get to the target output.  A series of additional exercises are structured in the same way. A target is presented. Then a walk-through of states is shown. A collection of slides that constitutes one wrangling exercise follows and a full set of exercises can be viewed here (https://tinyurl.com/y6b9z5vx). 


```{r eval=F, echo = F}
# xaringan html to pdf
pagedown::chrome_print("toy_exercise.html")

# converting pages of pdf pages to .png
# and slapping on a number
dir <- "figures"

if(!dir.exists(dir)) {dir.create(dir)}


for (i in 0:4){
  
magick::image_read_pdf("toy_exercise.pdf", pages = 8 + i) %>% 
  magick::image_border("grey15", "20x10") %>% # create images
  magick::image_annotate(text = i + 1, gravity = "center", 
                         size = 1600, color = "#00000033") %>% 
  magick::image_write(path = paste0("figures/", i,  "my_flipbook.png")) 
  
}
```

```{r, eval = T, echo = F, out.width="50%", fig.cap="Slides constituting one data wrangling task.",  fig.show='hold'}
dir <- "figures"

dir %>% 
  list.files(path = ., pattern = "my_") %>% 
  paste0(dir, "/", .) %>% 
knitr::include_graphics()
```


# What might the student/instructor experience look like?

Some students will be able to wrangle to accomplish the tasks just by inspecting the target.  In that case, students can build the data manipulation pipeline without looking ahead.  After producing the target they are encouraged to compare their strategy to that presented in the walk-through.  

Other students may initially be stumped about how to manipulate to the target. These students are encouraged to look ahead to the walk-through. They may be able to accomplish the task by advancing with the intermediary outputs as a guide.    

As an instructor, the task of assisting students who have trouble may be more manageable.  Students are less likely to overwhelmed by the data manipulation challenge given the step-by-step presentations.  Even if they are overwhelmed, the task is already logically broken up for the instructor or a fellow student to talk through the manipulation strategies.

Hopefully students will also come away with the idea:  *you may have some big task ahead of you, but stay cool.  Try to break up the task into pieces --- it will probably feel more manageable.*  


# Data manipulation as a general problem 

I think an advantage of this outcome-oriented exercise is that it de-emphasizes specific tools. Sometimes the focus of data manipulation exercises is syntactical. There is a lot of  focus on the tools being used.  But this exercise puts the data transformation itself center stage (subsetting, dropping columns, summarizing, etc.).  In fact, a large number of statistical software or syntactical paradigms could be used to accomplish the manipulation tasks presented here.  

I personally teach and built this exercise using 'the tidyverse' which does have some signatures in the output (like a separate group_by step when compared to data.table for example).  But the manipulation tasks could also be solved in Stata, SQL, SPSS, SAS, R tidyverse, base R, data.table, Python's Pandas -- or something yet to be invented!  This shifts the emphasis from a *how* question (specific implementation and syntax) to a more general *what* question. The question "what are the data manipulation moves that I can make?" is primary, and the how-to syntax is secondary. 

# How to build your own exercise

Educators should feel free to use this exercise as is or can create their own outcome-oriented tasks using the new {flipbookr} package (https://github.com/EvaMaeRey/flipbookr)! 

If you are familiar with R and Rmarkdown, you're in a good place to start building these outcome-oriented educational tasks.  Still, some things will be new and I'll discuss them here briefly.  For more detailed guidance it will also be useful to look at a basic tutorial (https://evamaerey.github.io/flipbooks/flipbook_recipes) on creating your first flipbooks and a short example (https://tinyurl.com/y25d2nhz) of exactly how to build an outcome-oriented exercise.

First, to get the movie-like experience of flipbooks, we need a slide show tool.  The `flipbookr` package is designed to be used with the Rmarkdown package `xaringan` to deliver material to .html slides.  Your Rmarkdown file will need to have the appropriate meta data set up so that these slides are produced.  
Once you have this in place, you can then write a code chunk with steps of a data manipulation task.  The code chunk options should be set to echo = FALSE and eval = TRUE so that the output displays but the code is hidden.  After a slide break, you'll then use the flipbookr function `chunk_reveal` to refer to the prepared chunk.  You should set `display_type` to "output" so that the output is shown but the code used to accomplish the manipulation is hidden.

````markdown
---

# Target #1

```{r task1, eval = FALSE}`r ''`
mtcars %>% 
  select(mpg:wt) %>% 
  arrange(-mpg) %>% 
  slice(1:5)
```
````

```markdown
---

``r 'r chunk_reveal(chunk_name = "task1", display_type = "output", title = "## Walk-through #1" )'``
``` 


You'll probably want to host the resulting html and associated files online for students.  Github Pages has worked well in my experience. 

# Going further

A bit more may be asked of students for futher engagement with the pipelines they have built.  At the end of data manipulation pipelines, you might ask students to create a descriptive name for an object that will hold the manipulated data or describe the data manipulation steps in words in groups. Also, educators might think about creating outcome oriented tasks like these for data visualization or data analysis!



```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 90%}
```



<!-- Chance is a quarterly non-technical statistics magazine published jointly by the American Statistical Association and Taylor \& Francis Group. It was established in 1988, and Taylor \& Francis has published it since 2012. The magazine sponsors the blog "The Statistics Forum", which allows anyone to post their thoughts on probability and statistics. -->

<!-- https://chance.amstat.org/ -->

<!-- https://chance.amstat.org/category/columns/taking-a-chance-in-the-classroom/ -->

<!-- --- -->

<!-- CHANCE is copublished quarterly by the American Statistical Association and Taylor & Francis Group The magazine is designed for anyone who has an interest in the analysis of data, informally highlighting sound statistical practice. CHANCE is not a technical magazine, but rather a cultural record of an evolving field, intended to entertain as well as inform. -->

<!-- Since its creation in 1988, CHANCE has covered such topics as the 1990 census adjustment and the redesigned population survey, sports, the environment, DNA evidence in the courts, a variety of medical issues—even how to win on "Jeopardy." CHANCE offers a unique opportunity to reach beyond statistics professionals to a more general audience. -->

<!-- While we invite authors to submit articles in ready-to-publish form, we also encourage prospective authors to submit ideas for articles or articles that are not yet in final form, and we will try to help with their development. -->

<!-- --- -->

<!-- Writing for CHANCE -->
<!-- We prefer short manuscripts of 10 to 15 double-spaced typewritten pages. Authors are encouraged to write in a relaxed, expository style; opinion, humor, and historical comments can (and should) be intermingled to make lively reading. Additional suggestions regarding writing for CHANCE are offered under Hints for Authors. -->

<!-- Submitting Files -->
<!-- Submit material via email as a PDF document and a Microsoft Word file to the editor at the following address: -->

<!-- Amanda Plunkett -->
<!-- CHANCE Editor -->
<!-- chancemag.editor@gmail.com -->
<!-- CHANCE Style -->
<!-- Details for listing and citing references are given in the ASA Magazine Style Guide (PDF). -->
