---
title: "Data manipulation practice"
author: "Evangeline Reynolds"
date: ""
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---






class: left, bottom, inverse
background-image: url(https://images.unsplash.com/photo-1533179856364-b2a4a66fa83e?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1267&q=80)
background-size: 87%
background-position: 40% 40%

# .column.Large[Wrangle]
## April 23, 2020
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>



#### .column[Gina Reynolds<br>  Photo Credit: Joshua Coleman]


```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}
```



```{r include = F}
knitr::opts_chunk$set(echo = F, cache = F, message = F, warning = F, comment = "")
library(flipbookr)
```

---
class: inverse, middle center

# Set up

---

#### The following exercises are designed to help you practice your data wrangling "vocabulary". In the exercises, you'll always start from the full gapminder data and wrangle to a target using functions from {dplyr} (one of the packages made available in the tidyverse).

--

### Load the following packages:

```{r, echo = T}
library(gapminder)  # The data package
library(tidyverse)  # Contains the data manipulation tools
```

--


#### Make sure to that the number of observations and columns in your result match the target outcome.  Let's recall the full dimensions of the gapminder data: 

```{r, echo = T}
# dim gives number of rows, then number of columns
dim(gapminder)
```

--

Some other features to have in mind:

- {gapminder} only reports data every 5 years, from 1952 to 2007
- only {142} countries are included in the data


---

## You might use some of the following functions from the {dplyr} package:

|function |action |
| :--- | ---: |
| filter()    | *keep rows (if true)*|
| select()    | *keep variables (or drop them `-`var)*|
| mutate()    | *create a new variable*|
| case_when() | *is used for "recoding" variable, often used with mutate()*|
| rename()    | *renaming variables*|
| arrange()   | *order rows based on a variable*|
| slice()     | *keep or drop rows based on row number |


---



# Target #1


```{r selection, echo = F}
gapminder %>% 
  filter(year == 2002) %>% 
  select(-continent, -lifeExp) %>% 
  mutate(gdp = gdpPercap * pop)
```



---

`r chunk_reveal("selection", display_type = "output", title = "### Walk through #1")`




---



# Target #2


```{r filterarrselect, echo = F}
gapminder %>% 
  filter(year == 2007) %>% 
  select(country, year, pop) %>% 
  mutate(pop_million = pop/ 1000000) %>% 
  filter(pop_million > 100) %>% 
  arrange(-pop) 
```


---

`r chunk_reveal("filterarrselect", display_type = "output", title = "### Walk through #2")`




---


# Target #3


```{r filterage, echo = F}
gapminder %>% 
  filter(year == 2007) %>% 
  select(country, lifeExp, year) %>% 
  filter(lifeExp >= 80) %>% 
  arrange(-lifeExp)
```


---

`r chunk_reveal("filterage", display_type = "output", title = "## Walk through #3")`




---
class: center, inverse, middle


# Congrats! You're done.
